---
title: "Final Project"
subtitle: "Data Science for Biologists, Spring 2021"
author: Dayna McKinney
output: 
  html_document:
    theme: cerulean
    highlight: pygments
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
# DON'T CHANGE THE NEXT LINE, except to add NEW options if you have any.
knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo = TRUE, include = TRUE)

# Load all libraries you use HERE:
library(tidyverse)
```

<br><br>

## Introduction

> This data set contains information on the World Cup forecasts from 2018. I chose this data set because I thought it would be interesting to investigate the statistics on the world cup season of 2018. In addition, this was a data set that stuck out to me that I thought I would take interest in learning more about.

[World cup forecasts](https://projects.fivethirtyeight.com/soccer-api/international/2018/wc_forecasts.csv)

```{r read-examine-data}
worldcup <- readr::read_csv('https://projects.fivethirtyeight.com/soccer-api/international/2018/wc_forecasts.csv')

dplyr::glimpse(worldcup)
```

<br><br>

## Question 1

#### How many groups made it to the quarter finals?

### Methods
```{r}
worldcup %>%
  select(make_quarters, group) %>%
  mutate(make_quarters = ifelse(
    make_quarters == 1, "yes", "no"),
    Quarterfinalist = make_quarters) %>%
  select(-make_quarters) -> worldcup1

ggplot(worldcup1) +
  aes(y=group, x=Quarterfinalist, color=group) +
  geom_point(size=3.0) +
  theme(legend.position = "none") +
  labs(x= "Quarter Finalists",
       y="Group",
    title="Group Quarter Finalists")
```
### Answer

Based on the data, 6 out of the 8 groups made it to the quarter finals.


<br><br>

## Question 2

#### How many groups made it to the semi finals?

### Methods
``` {r}
worldcup %>%
  select(make_semis, group) %>%
  mutate(make_semis = ifelse(
    make_semis == 1, "yes", "no"),
    Semifinalist = make_semis) %>%
  select(-make_semis) -> worldcup2

ggplot(worldcup2) +
  aes(y=group, x=Semifinalist, color=group) +
  geom_point(size=3.0) +
  theme(legend.position = "none") +
  labs(x= "Semi Finalists",
       y="Group",
    title="Group Semi Finalists")
  
```
### Answer

Based on the data, 3 out of the 8 groups made it to the semi finals.

<br><br>

## Question 3

#### How many groups made it to the finals?

### Methods
```{r}
worldcup %>%
  select(make_final, group) %>%
  mutate(make_final = ifelse(
    make_final == 1, "yes", "no"),
    Finalist = make_final) %>%
  select(-make_final) -> worldcup3

ggplot(worldcup3) +
  aes(y=group, x=Finalist, color=group) +
  geom_point(size=3.0) +
  theme(legend.position = "none") +
  labs(x= "Finalists",
       y="Group",
    title="Group Finalists")

```

### Answer

Based on the data, 2 out of the 8 groups made it to the finals. 


<br><br>

## Question 4

#### Which group won the league?

### Methods
```{r}
worldcup %>%
  select(win_league, group) %>%
  mutate(win_league = ifelse(
    win_league == 1, "yes", "no"),
    leaguewinner = win_league) %>%
  select(-win_league) -> worldcup4

ggplot(worldcup4) +
  aes(y=group, x=leaguewinner, color=group) +
  geom_point(size=3.0) +
  theme(legend.position = "none") +
  labs(x= "League Winner",
       y="Group",
    title="League winner")

```

### Answer

Based on the data, group C won the league.
